Impacts of Month, Grocery Vendor, Presale Prices, and Historical Pricing on Discounted Egg Dozen Prices

### 2024-12-13 paper/paper.qmd
added heading + formula centred
citaiton
added package citations

library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(modelsummary)
library(rstanarm)
library(knitr)
library(kableExtra)
library(jtools)

@@

In Toronto, a city that has an absurdly high cost of living at $1,492.10 per single person even without rent included, any area where unnecessary spending can be avoided would be greatly advantageous to know [@numbeo]. For those who are trying to save as much as they can, grocery shopping and homecooking is often a strategy employed to avoid the additional expenses of eating out. Even for people don't cook often, grocery shopping is still a monthly expense that is nearly unavoidable. Within these items, eggs are a staple grocery item that is consumed often by a wide demographic of people, with the average Canadian consuming up to 242 eggs per year [@eggfacts]. It is a cheaper source of protein than meats and a common ingredient in many dishes, making it one of the most frequently purchased grocery items.

@@

The estimand that is targetted within this paper's analysis is the price of a dozen eggs across the 8 top Canadian grocers: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods. In addition to vendor, the other variables that will also be taken into consideration are each item's presale price, month of the sale, and average historical discount pricing of egg dozens, as measured by the average discounted price of eggs in the month prior to the current sale.

Through building a Multiple Linear Regression Model, this paper finds that all four variables had non-negligable influences on the discounted price of eggs, with a key interaction term between historical pricing and presale pricing playing a role as well. Notably, T&T and April offering the most discounted egg prices comparative to all others in their own category. As well, presale price was found to have a negative relationship with discounted price, indicating that the most expensive eggs will be discounted more, when all other variables are held constant.

@@

Under the guidance of @tellingstories, the R programming language [@citeR] was used for analysis of this dataset. The package `tidyverse` [@tidyverse] was used to simulate data and test the simulated data before analysis. Packages `tidyverse` [@tidyverse] and `arrow` [@arrow] were utilized to clean the full raw dataset,  which were then also used alongside packages `testthat` [@testthat] and `here` [@here] to test the cleaned dataset. Lastly, packages `tidyverse` [@tidyverse], `rstanarm` [@rstanarm], `arrow` [@arrow], and `here` [@here] were used to build predictive models using the cleaned dataset.

@@

The observations in the dataset are scraped from the grocer's official website's interface, which means it only contains the information that is publically listed and available. The recorded prices for each item are the listed price for the "in store pick up" option, with the target pickup neighbourhood being a neighbourhood in Toronto. Certain values and listings may be missed, since the internal APIs that power the grocer's websites are not accessed and specific extracts may error for certain vendors on particular days for unforeseeable reasons. Starting from July 11, the targeted variety of grocery items for data collection was greatly increased, meaning that there may be products which are missing pricing information before then.

@@

@fig-monthsales shows the number of sales provided on egg dozens across the months, with the least number of sales occuring in July and most in November.

@@

@@

$Y_i = \beta_0 + \beta_1 month + \beta_2 vendor + \beta_3 old\_price * prev\_month\_avg$

We create the model in R [@citeR], with help from package `tidyverse` [@tidyverse].

@@

The first weakness of this model is that it assumes a linear relationship between the discounted price and the list of predictor variables. This oversimplication could be missing many aspects of their true relationship if it is nonlinear. As well, there could be real-world events such as economic changes or policy changes which could have significant impact on pricing, but this model would fail to capture these influences.

@@

#modelsummary(
#  list(
#    "Model Results" = model
#  ),
#  fmt = 2
#)

@@

The $R^2$ and Adjusted $R^2$ values of 0.82 indicate that 82% of the variability in the discounted price within this dataset can be explained by the current model.

@@

Frequency of sales occuring is also an important consideration because it impacts the accuracy of the model in predicting prices of eggs sold at that chain. That is, the observation that Galleria has very few sale events also has the effect that the model's accuracy when applied to Galleria listings will potentially be lower than when applied to Metro, which has a much higher number of discounts within this dataset.

>>>>

library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
#library(modelsummary)
#library(rstanarm)
library(knitr)
library(kableExtra)
library(jtools)

@@

In Toronto, a city that has an absurdly high cost of living at $1,492.10 per single person even without rent included, any area where unnecessary spending can be avoided would be greatly advantageous to know [@numbeo]. For those who are trying to save as much as they can, grocery shopping and home cooking is often a strategy employed to avoid the additional expenses of eating out. Even for people don't cook often, grocery shopping is still a monthly expense that is nearly unavoidable. Within these items, eggs are a staple grocery item that is consumed often by a wide demographic of people, with the average Canadian consuming up to 242 eggs per year [@eggfacts]. It is a cheaper source of protein than meats and a common ingredient in many dishes, making it one of the most frequently purchased grocery items.

@@

The estimand that is targeted within this paper's analysis is the price of a dozen eggs across the 8 top Canadian grocers: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods. In addition to vendor, the other variables that will also be taken into consideration are each item's presale price, month of the sale, and average historical discount pricing of egg dozens, as measured by the average discounted price of eggs in the month prior to the current sale.

Through building a Multiple Linear Regression Model, this paper finds that all four variables had non-negligible influences on the discounted price of eggs, with a key interaction term between historical pricing and presale pricing playing a role as well. Notably, T&T and April offering the most discounted egg prices comparative to all others in their own category. As well, presale price was found to have a negative relationship with discounted price, indicating that the most expensive eggs will be discounted more, when all other variables are held constant.

@@

Under the guidance of @tellingstories, the R programming language [@citeR] was used for analysis of this dataset. The package `tidyverse` [@tidyverse] was used to simulate data and test the simulated data before analysis. Packages `tidyverse` [@tidyverse] and `arrow` [@arrow] were utilized to clean the full raw dataset,  which were then also used alongside packages `testthat` [@testthat] and `here` [@here] to test the cleaned dataset. Lastly, packages `tidyverse` [@tidyverse], `rstanarm` [@rstanarm], `arrow` [@arrow], `modelsummary` [@modelsummary], and `here` [@here] were used for exploratory data analysis before creating the final model. The packages `arrow` [@arrow], `here` [@here], `ggplot2` [@ggplot], `tidyverse` [@tidyverse], `jtools` [@jtools], `knitr` [@knitr], and `kableExtra` [@kableExtra] were used for visualizing the dataset and model results below.

@@

The observations in the dataset are scraped from the grocer's official website's interface, which means it only contains the information that is publicly listed and available. The recorded prices for each item are the listed price for the "in store pick up" option, with the target pickup neighbourhood being a neighbourhood in Toronto. Certain values and listings may be missed, since the internal APIs that power the grocer's websites are not accessed and specific extracts may error for certain vendors on particular days for unforeseeable reasons. Starting from July 11, the targeted variety of grocery items for data collection was greatly increased, meaning that there may be products which are missing pricing information before then.

@@

@fig-monthsales shows the number of sales provided on egg dozens across the months, with the least number of sales occurring in July and most in November.

@@

### Model Set-Up {#sec-model-setup}

@@

$$
Y_i = \beta_0 + \beta_1 month + \beta_2 vendor + \beta_3 old\_price * prev\_month\_avg
$$

We create the model in R [@citeR], with help from package `tidyverse` [@tidyverse]. Model results are displayed with help from package `jtools` [@jtools].

@@

The first weakness of this model is that it assumes a linear relationship between the discounted price and the list of predictor variables. This oversimplification could be missing many aspects of their true relationship if it is nonlinear. As well, there could be real-world events such as economic changes or policy changes which could have significant impact on pricing, but this model would fail to capture these influences.

@@

@@

The $R^2$ and Adjusted $R^2$ values of 0.82 indicate that 82% of the variability in the discounted price within this dataset can be explained by the current model, which is strong evidence that these chosen variables do have influence on the estimand.

@@

Frequency of sales occurring is also an important consideration because it impacts the accuracy of the model in predicting prices of eggs sold at that chain. That is, the observation that Galleria has very few sale events also has the effect that the model's accuracy when applied to Galleria listings will potentially be lower than when applied to Metro, which has a much higher number of discounts within this dataset.
